A comparison of design-based and model-based approaches for finite population spatial sampling and inference.

### 2022-04-19 inst/manuscript/manuscript.tex
more minor updates

We studied performance of the four sampling-inference combinations by
selecting 2000 random SRS and GRTS samples of size \(n = 50\),
\(n = 100\), and \(n = 200\) from the realized ZMMI and Hg ppb
populations and then analyzing the samples using MB and DB inference. In
total, there were six separate scenarios (two responses crossed with
three sample sizes). Within each SRS and GRTS sample, design-based and
model-based inferences were used to estimate or predict the population
mean and construct 95\% coverage intervals. With model-based inference,
the exponential covariance was assumed, and covariance parameters and
the population mean were estimated (using REML) (separately for each SRS
and GRTS sample). We used the same evaluation metrics as for the
simulated data: mean bias, RMSE, and interval coverage. Mean bias is
taken as the average deviation between each sample's estimated (or
predicted) mean (\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of
ZMMI or Hg ppb): \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\),
where \(i\) indexes simulation trials. Root-mean-squared error is taken
as the square root of the average squared deviation between each
sample's estimated (or predicted) mean and its population mean:
\(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
Interval coverage is taken as the proportion of simulation trials where
the population mean is contained in its 95\% confidence (or prediction)
interval. These intervals are constructed using the normal distribution.

>>>>

We studied performance of the four sampling-inference combinations by
selecting 2000 random SRS and GRTS samples of size \(n = 50\),
\(n = 100\), and \(n = 200\) from the realized ZMMI and Hg ppb
populations and then analyzing the samples using MB and DB inference. In
total, there were six separate scenarios (two responses crossed with
three sample sizes). Within each SRS and GRTS sample, design-based and
model-based inferences were used to estimate or predict the population
mean and construct 95\% coverage intervals. With model-based inference,
the exponential covariance was assumed, and covariance parameters and
the population mean were estimated using REML (separately for each SRS
and GRTS sample). We used the same evaluation metrics as for the
simulated data: mean bias, RMSE, and interval coverage. Mean bias is
taken as the average deviation between each sample's estimated (or
predicted) mean (\(\hat{\mu}_i\)) and the population mean (\(\mu\)) (of
ZMMI or Hg ppb): \(\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)\),
where \(i\) indexes simulation trials. Root-mean-squared error is taken
as the square root of the average squared deviation between each
sample's estimated (or predicted) mean and its population mean:
\(\sqrt{\frac{1}{n}\sum_{i = 1}^{2000} (\hat{\mu}_i - \mu)^2}\).
Interval coverage is taken as the proportion of simulation trials where
the population mean is contained in its 95\% confidence (or prediction)
interval. These intervals are constructed using the normal distribution.
